---
title: "Beginning Exercise"
output: html_document


# 1. Generating 'Turkce' for Turkish and 'Matematik' for Math score vectors (0-100) for 20 students. 

```{r}
options(width = 100) # make the row width = 100
Turkce <- c(75, 80, 45, 90, 60, 70, 80, 50, 95, 65, 100, 100, 30, 85, 75, 65, 55, 90, 70, 80)
Matematik <- c(100, 70, 55, 80, 40, 85, 60, 55, 75, 45, 90, 90, 25, 75, 65, 50, 45, 80, 60, 70)
```

# 2. Printing the data

```{r}
print(Turkce)
print(Matematik)
```

# 3. Minimum, maximum, mean, standard deviation, median, variance

## For Turkish scores

```{r}
min_tr <- min(Turkce)
max_tr <- max(Turkce)
mean_tr <- mean(Turkce)
sd_tr <- sd(Turkce)
median_tr <- median(Turkce)
var_tr <- var(Turkce)
```

## For Math scores

```{r}
min_mat <- min(Matematik)
max_mat <- max(Matematik)
mean_mat <- mean(Matematik)
sd_mat <- sd(Matematik)
median_mat <- median(Matematik)
var_mat <- var(Matematik)
```

#  4. Comparing mean, standard deviation and variance of scores

## if the result is positive, mean of Turkish scores is higher than Math scores. 

```{r}
mean_compare <- mean_tr - mean_mat
print(mean_compare)
```

## if the result is positive, standard deviation of Turkish scores is higher than Math scores.

```{r}
sd_compare <- sd_tr - sd_mat
print(sd_compare)
```

## if the result is positive, variance of Turkish scores is higher than Math scores

```{r}
var_compare <- var_tr - var_mat
print(var_compare)
```



