---
title: "Data Visualization Basics"
output: html_document
---

# 1. Scatter Plots
Scatter plots are used to visualize the relationship between two numeric variables. In

measurement, we often use this to see the correlation between two different tests.

## plot()

The `plot()` function is the most basic way to create scatter plots or line graphs.

```{r}
# Midterm and Final scores of 6 students
midterm <- c(50, 65, 70, 85, 90, 45)
final <- c(55, 60, 75, 80, 95, 50)

# Create a scatter plot
plot(midterm, final, 
     main = "Relationship Between Midterm and Final",
     xlab = "Midterm Scores", 
     ylab = "Final Scores",
     col = "blue", 
     pch = 19) # pch defines the point shape
```

# 2. Histograms
A histogram shows the frequency distribution of a single variable. It is essential for

checking if test scores follow a normal distribution.

## hist()

hist() groups data into bins and shows how many observations fall into each bin.

```{r}
# Simulated scores for 100 students
set.seed(123)
all_scores <- rnorm(100, mean = 70, sd = 10)

# Create a histogram
hist(all_scores, 
     main = "Distribution of Exam Scores",
     xlab = "Scores", 
     col = "darkgreen", 
     border = "white")
```

# 3. Boxplots
Boxplots (box-and-whisker plots) provide a visual summary of the data, including the median, 

quartiles, and potential outliers.

## boxplot()

It is perfect for comparing different groups or identifying students who performed

significantly differently from the rest.

```{r}
# Scores of two different classes
class_A <- c(60, 70, 75, 80, 85, 90, 95, 100)
class_B <- c(40, 50, 55, 60, 65, 70, 75, 80)

# Create a boxplot comparing two groups
boxplot(class_A, class_B, 
        main = "Comparing Class A and Class B Performance",
        names = c("Class A", "Class B"),
        col = c("lightblue", "orange"),
        ylab = "Scores")
```