---
title: "Data Counting and Matching in R"
output: html_document
---

# 1. Basic Calculations and Dimensions

Before starting an analysis, we may need to know the total score or the size of our dataset.

## sum(), nrow(), and ncol()

sum() adds up all numeric values in a vector.

nrow() returns the number of rows (e.g., number of students).

ncol() returns the number of columns (e.g., number of variables or test items).

```{r}
scores <- c(10, 20, 30, 40)
# Total score
sum(scores)

# Creating a small data frame (example table)
df <- data.frame(ID = 1:5, Score = c(80, 90, 70, 85, 60))

# How many students (rows) are there?
nrow(df)

# How many variables (columns) are there?
ncol(df)
```

# 2. Finding Specific Indices

Sometimes we don't want the value itself, but its location in the list.

## which()

which() returns the positions (indices) of elements that meet a specific condition.

```{r}
exam_scores <- c(45, 70, 85, 90, 30, 85)

# Which indices have a score higher than 80?
which(exam_scores > 80)

# Which index is exactly 30?
which(exam_scores == 30)
```

# 3. Matching and Membership

These functions are used to compare two different lists (e.g., comparing a student list with a prize-winner list).

## %in% and match()

%in% checks if each element of the first list exists in the second list (returns TRUE/FALSE).

match() returns the position of the first occurrence of a value in another vector.

```{r}
all_students <- c("Alice", "Bob", "Charlie", "David")
winners <- c("Bob", "David")

# Is each student a winner? (Logical check)
all_students %in% winners

# Where is "Charlie" located in the student list?
match("Charlie", all_students)
```

