---
title: "Data Transformation and Ranking in R"
output: html_document
---

# 1. Data Type Conversion
In R, variables can be stored as different types (numeric, character, etc.). Sometimes we need to convert them, especially when importing data from external sources.

## as.character() and as.numeric()
`as.character()` converts a value to text, while `as.numeric()` converts it to a number. 

```{r}
# A score stored as a character
raw_score <- "85"
# Convert to numeric to perform calculations
numeric_score <- as.numeric(raw_score)

# A student ID stored as a number
student_no <- 101
# Convert to character for labeling
id_label <- as.character(student_no)

numeric_score * 2
```

# 2. Indexing and Selection
To analyze specific parts of our data, we use square brackets [].

```{r}
scores <- c(55, 70, 85, 90, 45)

# Select the 3rd student's score
scores[3]

# Select multiple scores (1st and 4th)
scores[c(1, 4)]
```

# 3. Finding Maximum Values
In assessment, we often need to identify the top performance. 

## max() and which.max()

max() gives the highest value, while which.max() tells you the position (index) of that value.

```{r}
exam_results <- c(65, 88, 92, 74, 95, 81)

# What is the highest score?
max(exam_results)

# Which student (by position) got the highest score?
which.max(exam_results)
```

# 4. Sorting and Ranking
Ordering data helps us see the performance distribution from lowest to highest or vice versa.

# sort(), order() and rank()

sort(): Returns the values in order.

order(): Returns the indices in the sorted order (useful for sorting whole tables).

rank(): Assigns a rank to each element (1st place, 2nd place, etc.).

```{r}
final_scores <- c(70, 95, 85, 60)

# Sort scores in ascending order
sort(final_scores)

# Get the indices to sort the scores
order(final_scores)

# Rank the students based on their scores
rank(final_scores)
```

